---
title: "OG GoaT File Comparison"
author: "Anu"
date: "`r Sys.Date()`"
output: html_notebook
---

## Input Files Required
This comparison script requires the following input files to be present in the same directory as this notebook:

1. `OG_species_goat_final.csv` → latest GoaT export  
2. `older_version.csv` → previous version for comparison  

Please ensure both files are in the current working directory before running the notebook.

# Step 1: Read both files (skip first 7 metadata rows)
```{r}
new_data <- read.csv(
  "OG_species_goat_final.csv",
  skip = 7,
  stringsAsFactors = FALSE
)

old_data <- read.csv(
  "older_version.csv",
  skip = 7,
  stringsAsFactors = FALSE
)

dim(new_data)
dim(old_data)
```

# Step 2: Check column names
```{r}
colnames(new_data)
colnames(old_data)
```

# Step 3: Compare row counts
```{r}
cat("New file rows:", nrow(new_data), "\n")
cat("Old file rows:", nrow(old_data), "\n")
cat("Difference:", nrow(new_data) - nrow(old_data), "\n")
```

# Step 4: Find newly added species
```{r}
new_species <- setdiff(
  new_data$species,
  old_data$species
)

new_species
length(new_species)
```

# Step 5: Find removed species
```{r}
removed_species <- setdiff(
  old_data$species,
  new_data$species
)

removed_species
length(removed_species)
```

# Step 6: Compare sequencing status changes
```{r}
library(dplyr)

status_changes <- old_data %>%
  select(species, sequencing_status_old = sequencing_status) %>%
  inner_join(
    new_data %>%
      select(species, sequencing_status_new = sequencing_status),
    by = "species"
  ) %>%
  filter(sequencing_status_old != sequencing_status_new)

status_changes
```

# Step 7: Compare target list status changes
```{r}
target_changes <- old_data %>%
  select(species, old_target = target_list_status) %>%
  inner_join(
    new_data %>%
      select(species, new_target = target_list_status),
    by = "species"
  ) %>%
  filter(old_target != new_target)

target_changes
```