---
title: "OG GoaT File Builder"
author: "Anushka S. Dissnaayaka M."
date: "2026-03-31"
output: html_notebook
---

## Script Information
- **Created by:** Anushka S. Dissnaayaka M.  
- **Created date:** 2026-03-31  
- **Last updated:** `r Sys.Date()`  
- **Purpose:** Build GoaT-compatible TSV and CSV files for Ocean Genomes species export

# Step 1: Read the exported CSV from DBeaver
```{r}
species_data <- read.csv(
  "species_goat_compatible.csv",
  stringsAsFactors = FALSE
)

head(species_data)
```

# Step 2: Create the 7 required header lines
```{r}
header_lines <- c(
  "# project_name\tOG",
  "# subproject_name\t",
  "# primary_contact\tShannon Corrigan",
  "# primary_contact_institution\tOcean Genomes Minderoo Ocean Omics Centre at University of Western Australia Crawley Perth Australia",
  "# primary_contact_email\toceangenomes@uwa.edu.au",
  paste0("# date_of_update\t", format(Sys.Date(), "%Y-%m-%d")),
  "# schema_version\tebp_species_goat_2.6"
)

header_lines
```

# Step 3: Create the 8-column header row
```{r}
column_header <- paste(colnames(species_data), collapse = "\t")

column_header
```

# Step 4: Convert all data rows into tab-separated lines
```{r}
data_lines <- apply(species_data, 1, function(x) {
  paste(x, collapse = "\t")
})

head(data_lines)
```

# Step 5: Combine header + column names + data
```{r}
final_output <- c(header_lines, column_header, data_lines)

head(final_output, 10)
```

# Step 6: Create output folder
```{r}
if (!dir.exists("output")) {
  dir.create("output")
}

dir.exists("output")
```

# Step 7: Save GoaT-ready TSV file (main final file)
```{r}
writeLines(
  final_output,
  "output/OG_species_goat.tsv"
)
```

# Step 8: Save a normal CSV for Excel review
```{r}
# Step 8: Save full GoaT file as CSV with 7 header rows
writeLines(
  gsub("\t", ",", final_output),
  "output/OG_species_goat.csv"
)
```
# Step 9: Check saved files
```{r}
list.files("output")
```